---
title: "M01-2-Introduction to Literate Programming-Application Assignment"
author: "Jonny De La Torre"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
format: 
  html: 
    toc: true
    toc-depth: 4
    number-sections: true
    code-line-numbers: true
    code-fold: false
    code-link: true
    embed-resources: true
editor: visual
execute: 
  freeze: auto
  warning: false
---

# Introduction

Literate programming combines code and documentation in a structured manner to improve readability and reproducibility. This assignment applies these principles to basic R programming exercises.

## Exercise 1: Create a Variable

```{r}
X <- "This is my first assignment"
```

## Exercise 2: Adding Texts in Base R

### (1) Learning `paste()` Function

```{r}
# Example usage of paste() and paste0()
paste("Hello", "World")
paste0("Hello", "World")
```

### (2) Using `paste()` to Append Text

```{r}
X <- paste(X, "and I'm loving it!")
X
```

## Exercise 3: Create and Modify a Vector

```{r}
Y <- c(2, 3, 4, 5)  # Creating the vector
Y <- Y * 2  # Multiplying by 2
```

## Exercise 4: Print Variables

```{r}
X  # Print variable X
Y  # Print vector Y
```

## Exercise 5: Find Max and Min of Y

```{r}
max(Y)  # Maximum value
min(Y)  # Minimum value
```

## Exercise 6: Load ggplot2 and View Data

```{r}
# Load ggplot2
library(ggplot2)

# View first six rows of economics dataset
head(economics)
```

## Exercise 7: Understanding and Plotting Data

### Definitions of Variables

-   **date**: The date of the observation.
-   **pce**: Personal consumption expenditures (in billions of dollars).
-   **pop**: Population in thousands.
-   **psavert**: Personal savings rate.
-   **uempmed**: Median duration of unemployment (in weeks).
-   **unemploy**: Total number of unemployed persons (in thousands).

### Scatter Plot of Two Related Variables

```{r}
plot <- ggplot(economics, aes(x = unemploy, y = uempmed)) +
  geom_point()

plot + 
  labs(title = "Unemployment vs. Median Duration",
       x = "Number of Unemployed Persons",
       y = "Median Duration of Unemployment (weeks)")
```

## Exercise 8: Replicating Plot with Pipe Operator

```{r}
library(dplyr)
library(ggplot2)

economics %>%
  ggplot(aes(x = unemploy, y = uempmed)) +
  geom_point() +
  labs(title = "Unemployment vs. Median Duration",
       x = "Number of Unemployed Persons",
       y = "Median Duration of Unemployment (weeks)")
```

## Exercise 9: Findings

::: callout-note
The scatter plot (Figure 1) demonstrates a positive correlation between the number of unemployed persons and the median duration of unemployment. As unemployment rises, people tend to remain unemployed for longer periods.
:::

## Exercise 10: Summary of Implemented Tools

1.  **Section numbering** using `#`, `##`, and `###`.
2.  **Bold font** via `**bold text**`.
3.  **Italicized font** via `*italicized text*`.
4.  **Callout-tip** for hints.
5.  **Callout-note** for findings.
6.  **Blockquotes** for explanations.
7.  **Tick marks** (\`\`\`) to highlight functions.
8.  **Cross-referencing** figures.
9.  **Piping (`%>%`)** for streamlined code.
10. **Formatted HTML output** with a table of contents and numbered sections.

# Conclusion

This assignment demonstrated the principles of literate programming using Quarto Markdown. By structuring code and explanations clearly, the document is both functional and readable for future reference.
